???A quantitative framework for investigating the reliability of network construction

### 2017-09-19 manuscript/CI_interactions.tex
fixed some small compiling errors

It is important to keep in mind that for network analysis, ecologists seek to measure the occurrence of true absences, which is the joint event $L=1\cup X=1 \cup \D=1$), but in reality they measure the marginal probability $P(L) = k/n$.

@@

As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur, Weinstein et al. (2017) used daily repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms (Weinstein, B. G., & Graham, C. H. (2017). On comparing traits and abundance for predicting species interactions with imperfect detection. Food Webs 11; 17-25: https://doi.org/10.1016/j.fooweb.2017.05.002). While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (like associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some rare interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, then it may still be either because it was impossible that day ((P(L|co-occurrence of A and B)=0), because interaction between the two did occur but was not observed, or from any combination of the two. Thus, from a conceptual perspective, this approach will fail to satisfactorily separating between all sources of uncertainty. An interaction not observed on day t may have gone unobserved either because there were constraints on the process (i.e. the probability was really zero), because of chance variation in the realized incidence of an event associated with a finite probability. Most importantly, if during a series of days, two species are simply never observed co-occurring, then we still have not learnt a thing about their probability of interacting should they do so. In other words, if the two interaction partners do not co-occur, we will still gain no information.

@@

The mean of these standardized degrees is 0.158, approximately half the $\theta_{MLE}$ obtained from the observed data. We can use these standardized degrees to inform our prior expectation of the 
distribution of interaction probabilities. With some simple R code (function calculate_parameters, \emph{Appendix S2}), we obtain prior parameters $\alpha$=0.7559544
and $\beta$=3.6710277. Using these priors in equations~\ref{mean} 
and~\ref{variance} above (or in the R function calculate_distribution in \emph{Appendix S2}), we find a prior mean interaction probability of 0.1707607 and a prior variance of 0.0319860.

@@

We may also wish to calculate a credible interval (analogous to the frequentist confidence interval). This is also quite straightforward (see function credible_interval in \emph{Appendix S2}). In this case, a 95\% credible interval for the mean of theta is (0.07755119, 0.50557509).

Now, consider the case where the two species have never been observed interacting across $n$ trials. The question is then ``what is the probability these two species do not interact''? Since it is not possible to prove that the two species could never interact, given the framework above we must fix a threshold below which we consider that there is no interaction. We call this threshold probability $\theta*$. We then use the cumulative distribution function to estimate $P(\theta<\theta*|X=0,n)$. The 
function plot_precision in \emph{Appendix S2} calculates distribution function for $\theta*$ with an increasing number of trials. The figure below illustrates this distribution for $\theta*$=0.1.

@@

      As well as obtaining posterior means and variances, it is possible to calculate a 95\% credible interval about the posterior mean. This is again straightforward to do in R (see the function "credible interval" in \emph{Appendix S2}). Calculating the 95\% credible intervals for species pairs with 0 to 50 observed co-occurrances without any observed interaction (as in Fig.~\ref{Salix_pdfs}), we see that the interval narrows dramatically as the number of observations increases. In no case, however, does the confidence interval include 0. Even for a pair of species which appeared together in all 374 sites but were never observed interacting, the confidence interval (0.001, 0.018) comes close to, but does not include zero.

      \begin{table}[h]
        \caption{}
        \label{interval_table}
        \begin{tabular}{l | c c}
        N & Lower bound & Upper bound \\
        \hline
        0   & 0.157 & 0.901 \\
        2   & 0.099 & 0.743 \\
        4   & 0.072 & 0.621 \\
        6   & 0.057 & 0.530 \\
        8   & 0.047 & 0.462 \\
        10 & 0.040 & 0.408 \\
        14 & 0.031 & 0.331 \\
        20 & 0.023 & 0.258 \\
        30 & 0.016 & 0.188 \\
        40 & 0.012 & 0.148 \\
        50 & 0.010 & 0.122 \\
        \hline
        374 & 0.001 & 0.018 \\
        \hline
        \end{tabular}
        \end{table}

@@

    Rather than calculating confidence intervals for a posterior distribution after collecting data, we may wish to know how many data points are necessary to obtain a given level of confidence that two co-occurring species do not interact. The number of samples needed will depend on both our desired level of confidence and the threshold below which we assume that two species are unlikely to ever interact. If we want to be 95\% confident that the interaction probability for two species in the \emph{Salix}-galler system is below 0.5, 0.1, or 0.01, we would need 6, 54, and 114 observed co-occurrances with no observed interaction (see function samples_required in~\emph{Appendix S2}). Given the low levels of replication in most network studies, this implies that we should have fairly low confidence in many "non-interacting" pairs of species. Even in the \emph{Salix}-galler dataset, [[how many pairs of species co-occur in >6,54,114 sites?]].  Discounting potential interactions, then, requires either a stronger prior expectation of no interaction (e.g., we can be highly confident that a palm tree will not predate upon a leopard no matter how many times the two species are observed together) or very extensive sampling.

>>>>

It is important to keep in mind that for network analysis, ecologists seek to measure the occurrence of true absences, which is the joint event $L=1\cup X=1 \cup D=1$), but in reality they measure the marginal probability $P(L) = k/n$.

@@

As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur, Weinstein et al. (2017) used daily repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms (Weinstein, B. G., \& Graham, C. H. (2017). On comparing traits and abundance for predicting species interactions with imperfect detection. Food Webs 11; 17-25: https://doi.org/10.1016/j.fooweb.2017.05.002). While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (like associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some rare interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, then it may still be either because it was impossible that day ((P(L|co-occurrence of A and B)=0), because interaction between the two did occur but was not observed, or from any combination of the two. Thus, from a conceptual perspective, this approach will fail to satisfactorily separating between all sources of uncertainty. An interaction not observed on day t may have gone unobserved either because there were constraints on the process (i.e. the probability was really zero), because of chance variation in the realized incidence of an event associated with a finite probability. Most importantly, if during a series of days, two species are simply never observed co-occurring, then we still have not learnt a thing about their probability of interacting should they do so. In other words, if the two interaction partners do not co-occur, we will still gain no information.

@@

The mean of these standardized degrees is 0.158, approximately half the $\theta_{MLE}$ obtained from the observed data. We can use these standardized degrees to inform our prior expectation of the 
distribution of interaction probabilities. With some simple R code (function calculate\_parameters, \emph{Appendix S2}), we obtain prior parameters $\alpha$=0.7559544
and $\beta$=3.6710277. Using these priors in equations~\ref{mean} 
and~\ref{variance} above (or in the R function calculate\_distribution in \emph{Appendix S2}), we find a prior mean interaction probability of 0.1707607 and a prior variance of 0.0319860.

@@

We may also wish to calculate a credible interval (analogous to the frequentist confidence interval). This is also quite straightforward (see function credible\_interval in \emph{Appendix S2}). In this case, a 95\% credible interval for the mean of theta is (0.07755119, 0.50557509).

Now, consider the case where the two species have never been observed interacting across $n$ trials. The question is then ``what is the probability these two species do not interact''? Since it is not possible to prove that the two species could never interact, given the framework above we must fix a threshold below which we consider that there is no interaction. We call this threshold probability $\theta*$. We then use the cumulative distribution function to estimate $P(\theta<\theta*|X=0,n)$. The 
function plot\_precision in \emph{Appendix S2} calculates distribution function for $\theta*$ with an increasing number of trials. The figure below illustrates this distribution for $\theta*$=0.1.

@@

      As well as obtaining posterior means and variances, it is possible to calculate a 95\% credible interval about the posterior mean. This is again straightforward to do in R (see the function "credible interval" in \emph{Appendix S2}). Calculating the 95\% credible intervals for species pairs with 0 to 50 observed co-occurrances without any observed interaction (as in Fig.~\ref{Salix_pdfs}), we see that the interval narrows dramatically as the number of observations increases. In no case, however, does the confidence interval include 0. Even for a pair of species which appeared together in all 374 sites but were never observed interacting, the confidence interval (0.001, 0.018) comes close to, but does not include zero (Table~\ref{interval_table}).

      \begin{table}[h]
        \caption{95\% credible intervals for the interaction probability of two species that co-occur in $N$ sites but are never observed to interact, obtained using prior information from the \emph{Salix}-galler network in~\citet{Balbour2016}.}
        \label{interval_table}
        \begin{tabular}{l | c c}
        N & Lower bound & Upper bound \\
        \hline
        0   & 0.157 & 0.901 \\
        2   & 0.099 & 0.743 \\
        4   & 0.072 & 0.621 \\
        6   & 0.057 & 0.530 \\
        8   & 0.047 & 0.462 \\
        10 & 0.040 & 0.408 \\
        14 & 0.031 & 0.331 \\
        20 & 0.023 & 0.258 \\
        30 & 0.016 & 0.188 \\
        40 & 0.012 & 0.148 \\
        50 & 0.010 & 0.122 \\
        \hline
        374 & 0.001 & 0.018 \\
        \hline
        \end{tabular}
        \end{table}

@@

    Rather than calculating confidence intervals for a posterior distribution after collecting data, we may wish to know how many data points are necessary to obtain a given level of confidence that two co-occurring species do not interact. The number of samples needed will depend on both our desired level of confidence and the threshold below which we assume that two species are unlikely to ever interact. If we want to be 95\% confident that the interaction probability for two species in the \emph{Salix}-galler system is below 0.5, 0.1, or 0.01, we would need 6, 54, and 114 observed co-occurrances with no observed interaction (Fig.~\ref{Salix_cdfs}; see function samples\_required in~\emph{Appendix S2}). Given the low levels of replication in most network studies, this implies that we should have fairly low confidence in many "non-interacting" pairs of species. Even in the \emph{Salix}-galler dataset, [[how many pairs of species co-occur in \textgreater6,54,114 sites?]].  Discounting potential interactions, then, requires either a stronger prior expectation of no interaction (e.g., we can be highly confident that a palm tree will not predate upon a leopard no matter how many times the two species are observed together) or very extensive sampling.
